---
title: 'Lab #4 (model)'
author: "Jerid Francom"
date: "9/12/2021"
output: 
  html_document: 
    toc: yes
    number_sections: yes
    df_print: kable
  pdf_document: 
    toc: yes
    number_sections: yes
    df_print: kable
bibliography: references.bib
---

```{r setup, include=FALSE}
library(tidyverse) # to manipulate and visualize datasets
library(janitor) # to create (cross-)tabulations
library(skimr) # to provide statistical summaries
```

# Read and inspect

```{r cedel-read-rds, include=FALSE}
cedel <- read_rds(file = "data/derived/cedel_dataset.rds") # read cedel_dataset.rds

glimpse(cedel) # overview of the dataset
```

In the CEDEL2 [@Lozano2009] dataset there are 1,931 compositions and seven variables. The variables `filename`, `sex`, and `proficiency` are categorical variables and the variables `placement_test_score_percent`, `num_tokens`, `num_types`, and `ttr` are continuous variables.

-   `filename` = The record number for each composition
-   `sex` = The biological sex of each L2 Spanish learner
-   `proficiency` = A standardized learner level associated with each learner
-   `placement_test_score_percent` = The score the learner received on a standardized placement test.
-   `num_tokens` = The number of words in each composition
-   `num_types` = The number of unique words in each composition.
-   `ttr` = The Type/ Token Ratio for each composition

# Single variables

**Categorical**

A summary of the categorical variable `sex` suggests that there are more females in this dataset than males.

```{r cedel-sex}
cedel %>% # dataset
  tabyl(sex) # create tabulation fo the `sex` variable
```

A summary of the categorical variable `proficiency` suggests that most learner levels have a similar number of participants, except for the lowest level ("Lower beginner"). 

```{r cedel-proficiency}
cedel %>% # dataset
  tabyl(proficiency) %>%  # create tabulation fo the `proficiency` variable
  arrange(desc(n)) # sort dataset by n in descending order
```

**Continuous**

Let's first look at a set of descriptive statistics for all the continuous variables. 

```{r cedel-continuous-variables-stats}
cedel %>% # dataset
  skim() %>% # get variety of numeric summaries
  yank("numeric") # select only continuous variables (type: numeric)
```

**The small histogram that appears when using the `skim()` function on numeric variables will cause issues when knitting to a PDF. To avoid them change the front matter under `pdf_document:` to include `latex_engine: xelatax`.**


The `ttr` appears to the show the least variability. 

I'll select `placement_test_score_percent` and `ttr` to summarize visually. Since they are single variables, a density plot will give me an understanding as to how these variables are distributed. 

```{r cedel-continuous-variables-plot}
p1 <- 
  cedel %>% # dataset
  ggplot(aes(x = placement_test_score_percent)) + # map placement_test_score_percent to x
  geom_density() # plot density

p2 <- 
  cedel %>% # dataset
  ggplot(aes(x = ttr)) + # map ttr to x
  geom_density() # plot density

gridExtra::grid.arrange(p1, p2, ncol = 2) # arrange the plots side-by-side
```

`placement_test_score_percent` is clearly non-normal, and may even be bi-model (having two modes, distinct clumping of groups). `ttr` does have a shape which visually more normal but to verify, I will run a normality test, in this case the Shapiro-Wilk normality test.


```{r cedel-shapiro-ttr}
shapiro.test(cedel$ttr) # check for normality
```

# Multiple variables

I would like explore two relationships: 

1. What is the relationship between a learner's placement score and their proficiency level? 

Since the `placement_test_score_percent` is continuous and `proficiency` is a categorical (ordinal) variable, I will use a box plot to visualize the relationship. 

```{r cedel-placement-proficiency}
cedel %>% # dataset
  ggplot(aes(x = proficiency, y = placement_test_score_percent)) + # map proficiency to x, and placement_test_score_percent to y
  geom_boxplot() + # create box plot
  labs(x = "Learner Proficiency Level", y = "Placement Test Score")
```
From this visual summary, it appears that placement test scores and proficiency are related. A small, but potentially interesting observation is the potentially larger variability of placement scores for learners in the 'Upper beginner', 'Lower intermediate', and 'Upper intermediate' levels.

2. What is the relationship between placement scores and lexical diversity (as measured by Type/ Token Ratio scores)?

Since both the `placement_test_score_percent` and `ttr` are continuous variables I will summarize this relationship with a scatter plot. I will also add a linear trendline. 

```{r cedel-placement-ttr, message=FALSE}
cedel %>% # dataset
  ggplot(aes(x = placement_test_score_percent, y = ttr)) + # map placement_test_score_percent to x and ttr to y
  geom_point() + # plot x/y points
  geom_smooth(method = "lm") + # plot linear trend line
  labs(x = "Placement Test Scores", y = "Type/Token Ratio")
```

From this plot there appears to be a relationship towards more lexical diversity for more advanced learner levels --but not a particularly strong one. Let's assess the strength of this correlation. As both variables are non-normally distributed, I will use "kendall" for the method.

```{r cedel-placement-ttr-correlation}
cor(x = cedel$placement_test_score_percent, y = cedel$ttr, method = "kendall")
```

Yes, the correlation coefficient is on the weaker side. 

# Assessment

...

# References
