---
title: "Curate CEDEL2 dataset"
author: "Jerid Francom"
date: "9/14/2021"
output: 
  html_document: 
    number_sections: yes
    df_print: kable
---

```{r setup, include=FALSE}
library(tidyverse) # for general data manipulation
library(janitor) # to normalize variable names
library(tidytext) # to tokenize text
```

Read the CEDEL2 data. Description of the process is documented in the Text as Data coursebook in Chapter 5 "Acquire data". 

```{r cedel-read-tsv}
cedel_org <- 
  read_tsv(file = "original/cedel2.tsv") %>% 
  clean_names()
```

Look at the data original `cedel_org` dataset. 

```{r cedel-glimpse}
glimpse(cedel_org)
```

Let's look a some descriptives and see which variables have complete cases (complete_rate). 

```{r cedel-skim}
skimr::skim(cedel_org)
```

OK. It looks like we can focus on a few variables to select: 

```{r}
cedel <- 
  cedel_org %>% 
  select(filename, sex, proficiency, placement_test_score_percent, text)

glimpse(cedel)
```

Let's tokenize the text and calculate some text measures (`num_tokens`, `num_types`, and `ttr`). 

```{r cedel-unnest-tokens-measures}
cedel <- 
  cedel %>% 
  unnest_tokens(output = words, input = text) %>% 
  group_by(filename) %>% 
  mutate(num_tokens = n(),
         num_types = n_distinct(words),
         ttr = round(num_types/ num_tokens, 3)) %>% 
  ungroup() %>% 
  distinct(filename, sex, proficiency, placement_test_score_percent, num_tokens, num_types, ttr)

glimpse(cedel)
```

Create factors. 

```{r cedel-factors-arrange}
cedel <- 
  cedel %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(proficiency = factor(proficiency, levels = c("Lower beginner", "Upper beginner", "Lower intermediate", "Upper intermediate", "Lower advanced", "Upper advanced")), ordered = TRUE) %>% 
  arrange(proficiency, sex)

cedel
```
Explore this curated dataset as an .rds file to maintain the factors in the dataset.

```{r cedel-write}
write_rds(x = cedel, file = "derived/cedel_dataset.rds")
```



